DMRGenie: Graphical interface for entanglement renormalization

### 2025-12-03 paper/paper.tex
Update paper.tex

In any scientific endeavour, the need to use numerical techniques is highly necessary. And for new advances to be made, new methods must be able to be implemented into any given language to evolve beyond the state-of-the-art. For anyone developing new algorithms, the need for a programming framework that enables as many solutions of the two-language problem as possible is necessary, making a Julia a strong contender for algorithmic development.

%For any algorithm development team, the need for custom software that displays as many solutions of the two-language problem is necessary, making Julia a strong option for use in algorithm development.

@@

Methods of entanglement renormalization use quantities native to quantum information ({\it i.e.}, the density matrix or the entropy) in order to optimize a a quantum problem to find solutions to the eigenvalue problem (Schr\"odinger's equation) \cite{townsend2000modern}. Entanglement renormalization techniques are formulated on a graph of tensors and have definite rules for tensor operations. Because tensor computations and operations can be cumbersome to represent mathematically in index notation, the field has adopted a diagram notation that allows for the improved communication of algorithms. Namely, the rank of the tensor corresponds to the number of lines on the diagram \cite{penrose1971angular}. An example is shown for rank-1, 2, and 3 tensors in Fig.~\ref{diagram}.

@@

When contracting two tensors together, one can connect two indices on the same tensor ({\it i.e.}, tracing over indices) or between tensors (contraction). When representing a network of tensors, the indices are typically not kept track of. Instead, indices that point towards each other or have matching labels are assumed to be contracted eventually. This rarely causes confusion when implementing algorithms.

@@

The use of the diagram notation is useful for representing large numbers of tensors in a coherent way without representing cumbersome mathematics (see Fig.~\ref{network}). For example, the wavefunction when decomposed through a series of reshapes and singular value decompositions (SVD) takes the form \cite{bakerCJP21}
\begin{align}\nonumber
|\psi\rangle&=\sum_{\substack{\sigma_1\sigma_2\sigma_3\\\sigma_4\sigma_5\sigma_6}}\sum_{\substack{a_1a_2a_3\\a_4a_5}}A^{\sigma_1}_{a_1}A^{\sigma_2}_{a_1a_2}A^{\sigma_3}_{a_2a_3}A^{\sigma_4}_{a_3a_4}A^{\sigma_5}_{a_4a_5}A^{\sigma_6}_{a_5}\\
&\quad\quad\quad\quad\quad\quad\quad\quad\times|\sigma_1\sigma_2\sigma_3\sigma_4\sigma_5\sigma_6\rangle
\label{mpsform}
\end{align}
which is far more intricate than the diagram notation but conveys no extra information. Any quantum operator (for example, the Hamiltonian) can be decomposed into a similar site-by-site form involving many tensors.

@@

We will first discuss the background software (DMRjulia and TENPACK; available in Julia's package system as {\tt DMRJtensor} and {\tt TensorPACK}) originally developed in 2016 and first introduced to Github in 2020 and full debut on the Julia package environment in 2021. This means that the library has spanned many version of the Julia version starting originally in v0.5 and now continuing to v1.11 and beyond. We believe that some choices from our implementation may be interesting to the broader community, although many of the improvements that were introduced here later appeared in this development independently appeared in Julia from other groups, showing that common issues are being encountered and handled in similar ways.

@@

The implementation here for methods of entanglement renormalization is given as 100\% of Julia code with no dependencies beyond the basic library.\footnote{As of this article, the library is in an extended beta-testing phase. As we add more users, we are creating useful error checks for algorithms that can be created. The library will be moved to the alpha phase after this is complete, and we welcome feedback and development from anyone even if this phase has passed.}

>>>>

In any scientific endeavour, the need to use numerical techniques is highly necessary. And for new advances to be made, new methods must be able to be implemented into any given language to evolve beyond the state-of-the-art. For anyone developing new algorithms, the need for a programming framework that enables as many solutions of the two-language problem as possible is necessary, making Julia a strong contender for algorithmic development.

@@

Methods of entanglement renormalization use quantities native to quantum information ({\it i.e.}, the density matrix or the entropy) in order to optimize a a quantum problem to find solutions to the eigenvalue problem \cite{townsend2000modern}. Entanglement renormalization techniques are formulated on a graph of tensors and have definite rules for tensor operations. Because tensor computations and operations can be cumbersome to represent mathematically in index notation, the field has adopted a diagram notation that allows for the improved communication of algorithms. Namely, the rank of the tensor corresponds to the number of lines on the diagram \cite{penrose1971angular}. An example is shown for rank-1, 2, and 3 tensors in Fig.~\ref{diagram}.

@@

%When contracting two tensors together, one can connect two indices on the same tensor ({\it i.e.}, tracing over indices) or between tensors (contraction). When representing a network of tensors, the indices are typically not kept track of. Instead, indices that point towards each other or have matching labels are assumed to be contracted eventually. This rarely causes confusion when implementing algorithms.

@@

The use of the diagram notation is useful for representing large numbers of tensors in a coherent way without representing the full mathematical detail (see Fig.~\ref{network}). For example, the wavefunction when decomposed through a series of reshapes and singular value decompositions (SVD) takes the form \cite{bakerCJP21}
\begin{align}\nonumber
|\psi\rangle&=\sum_{\substack{\sigma_1\sigma_2\sigma_3\\\sigma_4\sigma_5\sigma_6}}\sum_{\substack{a_1a_2a_3\\a_4a_5}}A^{\sigma_1}_{a_1}A^{\sigma_2}_{a_1a_2}A^{\sigma_3}_{a_2a_3}A^{\sigma_4}_{a_3a_4}A^{\sigma_5}_{a_4a_5}A^{\sigma_6}_{a_5}\\
&\quad\quad\quad\quad\quad\quad\quad\quad\times|\sigma_1\sigma_2\sigma_3\sigma_4\sigma_5\sigma_6\rangle
\label{mpsform}
\end{align}
which is far more intricate than the diagram notation but conveys no extra information. Any quantum operator (for example, the Hamiltonian) can be decomposed into a similar site-by-site form involving many tensors.

@@

We will first discuss the background software (DMRjulia and TENPACK; available in Julia's package system as {\tt DMRJtensor} and {\tt TensorPACK}) originally developed in 2016 and first introduced to Github in 2020 and full debut on the Julia package environment in 2021. This means that the library has spanned many version of the Julia version starting originally in v0.5 and now continuing to v1.11 and beyond. We believe that some choices from our implementation may be interesting to the broader community, although many of the improvements that were introduced here also independently appeared in Julia from other groups, showing that common issues are being encountered and handled in similar ways.

@@

%The implementation here for methods of entanglement renormalization is given as 100\% of Julia code with no dependencies beyond the basic library.\footnote{As of this article, the library is in an extended beta-testing phase. As we add more users, we are creating useful error checks for algorithms that can be created. The library will be moved to the alpha phase after this is complete, and we welcome feedback and development from anyone even if this phase has passed.}
